Variational Inference for Inverse Reinforcement Learning with Gaussian Processes

### 2018-11-22 mprop/mprop.tex
A bit minimalistic, but readable

Given this model, data $\mathcal{D}$, and inducing feature matrix
$\mathbf{X_u}$, our goal is then to find optimal values of hyperparameters
$\bm\lambda$, inducing rewards $\mathbf{u}$, and the rewards for all relevant
states $\mathbf{r}$. While the previous paper to consider this IRL model
computed maximum likelihood estimates for $\bm\lambda$ and $\mathbf{u}$, and
made an assumption that $\mathbf{r}$ in \eqref{eq:r} has zero variance
\cite{DBLP:conf/nips/LevinePK11}, we aim to avoid this assumption and use
VI to approximate the full posterior distribution $\posterior$.
\emph{Variational inference} is an approximation technique for probability
densities \cite{blei2017variational}. Let $\approximation$ be our approximating
family of probability distributions for $\posterior$ with its own hyperparameter
vector $\bm\nu$. Then the job of VI algorithms is to optimise $\bm\nu$ in order
to minimise the \emph{Kullback-Leibler} (KL) divergence between the original
probability distribution and our approximation.  KL divergence (asymmetrically)
measures how different the two distributions are, and in this case can be
defined as \cite{blei2017variational}
\[ \begin{split}
    \DKL(\approximation || \posterior) &= \Eq [\log\approximation -
    \log\posterior ] \\
    &= \Eq[\log\approximation - \log\pfull] \\
    &+ \Eq[\log p(\mathcal{D}, \mathbf{X_u})].
  \end{split}
\]
The last term is both hard to compute and constant w.r.t. $\approximation$
\cite{blei2017variational}, so we can remove it from our optimisation objective.
The negation of what remains is often called the \emph{evidence lower bound}
(ELBO) and is defined as\footnote{Throughout the proposal, all integrals should
  be interpreted as definite integrals over the entire sample space.}
\cite{DBLP:books/lib/Bishop07,blei2017variational}
\begin{equation} \label{eq:elbo}
  \begin{split}
    \mathcal{L}(\bm\nu) &= \Eq \left[ \log \frac{\pfull}{\approximation}
    \right] \\
    &= \iiint \approximation \log
    \frac{\pfull}{\approximation}\,d\bm\lambda\,d\mathbf{u}\,d\mathbf{r}.
  \end{split}
\end{equation}

@@

Levine et al. \cite{DBLP:conf/nips/LevinePK11} are the first to suggest a way to
learn nonlinear reward functions without harsh restrictions on the problem
domain by using GPs. We base our work primarily on their paper, and the
weaknesses we hope to address have already been covered in the previous section.
A recent extension to their work by Jin et al. \cite{DBLP:conf/uai/JinDAS17}
aims to harness the power of deep learning by using several layers of GPs,
making the model less dependent on being provided good features. They also use
VI, but with a few simplifying assumptions: deterministic training conditional
for the reward vector, and fully independent training conditional for the latent
state (see Section \ref{sec:vi_literature} for more details).

@@

Table \ref{table:approximations} further summarises some of the recent and/or
influential GP approximation approaches that might be relevant to our situation.
For many of the variables in our model (including the inducing points
located somewhere in $\mathbb{R}^d$ and the hyperparameters of the covariance
function), we have three possible ways of handling them:
\begin{itemize}
\item Provide a full Bayesian treatment by defining a prior probability
  distribution. While this option can easily prevent overfitting, the choice of
  prior can be difficult to justify.
\item Fix the value of the variable. We can then optimise the value outside of
  the main algorithm, e.g., by running the algorithm with a set of possible
  values and keeping the value that produces the best performance. This option
  can be slow if we are interested in finding the optimal value, but could be
  appropriate if the variable has little impact on the algorithm's performance.
\item Treat the variable as a variational parameter. This method will
  efficiently optimise the value of the variable, but is vulnerable to
  overfitting. Furthermore, this requires us to find the derivative of the ELBO
  w.r.t. the variable, which could be difficult.
\end{itemize}

@@

% TODO: the real reason is that we want to use the prior

In this section we show a feasible way to apply VI to the problem. Section
\ref{sec:preliminaries} defines several previously-missing parts of the model
and provides intuition for how the approximating distribution could be
structured. Section \ref{sec:structure} then defines the approximating
distribution, where some choices are justified by a body of literature, and some
are novel. Next, Section \ref{sec:elbo} derives a simplified expression for the
ELBO. Finally, in Section \ref{vi_algs} we describe our strategy for
optimising the ELBO.

@@

In order to properly investigate the difference between variational inference
and maximum likelihood estimation for the model, we keep other parts of the
model the same. Namely, we set the covariance function to a version of the
automatic relevance detection kernel
\cite{DBLP:conf/nips/LevinePK11,neal2012bayesian}
\[ k_{\bm\lambda}(\mathbf{x}_i, \mathbf{x}_j) = \lambda_0\exp\left(
    -\frac{1}{2}(\mathbf{x}_i - \mathbf{x}_j)^\intercal\bm\Lambda(\mathbf{x}_i -
    \mathbf{x}_j) - \mathbbm{1}[i \ne j]\sigma^2\Tr[\bm\Lambda] \right), \]
where $\lambda_0$ is the overall `scale' factor for how similar or distant the
states are, $\bm\Lambda = \diag(\lambda_1, \dots, \lambda_d)$ is a diagonal
matrix that determines how relevant each feature is (where $d$ denotes the
number of features), $\mathbbm{1}$ is defined as
\[ \mathbbm{1}[b] = \begin{cases}
    1 & \mbox{if $b$ is true} \\
    0 & \mbox{otherwise},
  \end{cases} \]
and $\sigma^2$ is set to $10^{-2}/2$ (as the original paper noted that
the value makes little difference to the performance of the algorithm
\cite{DBLP:conf/nips/LevinePK11}). Our vector of hyperparameters for the
covariance function is then $\bm\lambda = (\lambda_0, \dots,
\lambda_d)^\intercal$. Similarly, we keep the expression for the prior of
$\bm\lambda$:
\begin{equation} \label{eq:theta_prior}
  p(\bm\lambda | \mathbf{X_u}) = \exp\left( -\frac{1}{2}\Tr[\Kuu^{-2}] -
    \sum_{i=1}^d \log(\lambda_i + 1) \right).
\end{equation}

Next, we can rewrite the posterior by using the chain rule and Bayes' theorem in
order to get a better sense of what we are trying to approximate:
\[
  \begin{split}
    \posterior &= p(\bm\lambda | \mathbf{X_u}, \mathcal{D}) \times p(\mathbf{u}
    | \bm\lambda, \mathbf{X_u}, \mathcal{D}) \times p(\mathbf{r} | \bm\lambda,
    \mathbf{X_u}, \mathbf{u}, \mathcal{D}) \\
    &\propto p(\bm\lambda | \mathbf{X_u}, \mathcal{D}) \times p(\mathbf{u} |
    \bm\lambda, \mathbf{X_u}, \mathcal{D}) \times p(\mathcal{D} | r) \times
    p(\mathbf{r} | \bm\lambda, \mathbf{X_u}, \mathbf{u}) \\
    &\propto p(\bm\lambda | \mathbf{X_u}, \mathcal{D}) \times p(\mathcal{D} |
    \bm\lambda, \mathbf{X_u}, \mathbf{u}) \times p(\mathbf{u} | \bm\lambda,
    \mathbf{X_u}) \times p(\mathcal{D} | r) \times p(\mathbf{r} | \bm\lambda,
    \mathbf{X_u}, \mathbf{u}) \\
    &\propto p(\mathcal{D} | \bm\lambda, \mathbf{X_u}) \times p(\bm\lambda |
    \mathbf{X_u}) \times p(\mathcal{D} | \bm\lambda, \mathbf{X_u}, \mathbf{u})
    \times p(\mathbf{u} | \bm\lambda, \mathbf{X_u}) \times p(\mathcal{D} | r)
    \times p(\mathbf{r} | \bm\lambda, \mathbf{X_u}, \mathbf{u})
  \end{split}
\]
Note that now there are only two unknown probability distributions:
$p(\mathcal{D} | \bm\lambda, \mathbf{X_u})$ and $p(\mathcal{D} | \bm\lambda,
\mathbf{X_u}, \mathbf{u})$, which can be expressed as follows:
\begin{align*}
  p(\mathcal{D} | \bm\lambda, \mathbf{X_u}, \mathbf{u}) = \int &p(\mathcal{D} | r) \times p(\mathbf{r} | \bm\lambda, \mathbf{X_u}, \mathbf{u}) \, d\mathbf{r}, \\
  p(\mathcal{D} | \bm\lambda, \mathbf{X_u}) = \iint &p(\mathcal{D} | r) \times p(\mathbf{r} | \bm\lambda, \mathbf{X_u}, \mathbf{u}) \times p(\mathbf{u} | \bm\lambda, \mathbf{X_u}) \, d\mathbf{u} \, d\mathbf{r}.
\end{align*}
This suggests the following form for the approximation:
\begin{equation} \label{eq:approximation}
  \approximation = q(\bm\lambda) \times q(\mathbf{u} | \bm\lambda) \times q(\mathbf{r}
  | \bm\lambda, \mathbf{u}).
\end{equation}

@@

At this point we are forced to make assumptions about the approximate posterior
in order to arrive at an implementable solution. Time permitting, ways to relax
the assumptions may be investigated towards the end of the project.

First, as is common in the literature for applying VI to GPs
\cite{DBLP:conf/nips/ChengB17,DBLP:conf/uai/HensmanFL13,DBLP:conf/icml/HoangHL15,DBLP:journals/jmlr/Titsias09},
we simply set
\begin{equation}
  q(\mathbf{r} | \bm\lambda, \mathbf{u}) = p(\mathbf{r} | \bm\lambda, \mathbf{X_u},
  \mathbf{u}).
\end{equation}
We can make a similarly justified choice for $q(\mathbf{u} | \bm\lambda)$:
\begin{equation}
  q(\mathbf{u} | \bm\lambda) = q(\mathbf{u}) = \mathcal{N}(\mathbf{u}; \mathbf{m}, \mathbf{S})
\end{equation}
for some variational parameters $\mathbf{m} \in \mathbb{R}^{m}$ and a positive
semi-definite matrix $\mathbf{S} \in \mathbb{R}^{m \times m}$.
\cite{DBLP:conf/nips/ChengB17,DBLP:journals/jmlr/HensmanDS17,DBLP:conf/aaai/HoangHL17}.

In order to have a reasonable way of calculating/approximating the ELBO (see
Section \ref{vi_algs}), we need to have an approximating distribution for
$\bm\lambda$, but unfortunately all the papers in Table \ref{table:approximations}
either fix it or treat it as a variational parameter. Hence we make our first
assumption without justification from previous literature:
\begin{equation} \label{eq:theta}
  q(\bm\lambda) = \prod_{i=0}^dq(\lambda_i).
\end{equation}
We want to restrict $\lambda_0$ to be positive so that $k_{\bm\lambda}$ would
produce non-negative values and not become trivial. Similarly, we want that
$\lambda_i \ge 0$ for $i = 1, \dots, d$ so that $\bm\Lambda$ is a
positive-definite matrix. Considering the possible distributions for all $d + 1$
variables, we would like the mean to be flexible (i.e., not tied to zero, like
with the exponential distribution), and the tails to converge to zero as the
value of the random variable moves away from the mean. We might want to support
some right skew, but the distribution should be close to symmetric with at least
some parameter values. This limits our choice of distributions quite
significantly, and we decide to go with the gamma distribution as it is fairly
flexible and commonly used \cite{hogg2018introduction}. We then define the
probability density functions as
\begin{equation} \label{eq:zeta}
  q(\lambda_i) = \Gamma(\lambda_i; \alpha_i, \beta_i) =
  \frac{\beta_i^{\alpha_i}}{\Gamma(\alpha_i)}\lambda_i^{\alpha_i - 1}e^{-\beta_i\lambda_i}, \quad i = 0, \dots, d,
\end{equation}
where $\alpha_i > 0$ and $\beta_i > 0$ are parameters of the distribution, and
$\Gamma(\cdot)$ is the gamma function. This gives us our
vector of variational parameters $\bm\nu = (\mathbf{m}, \mathbf{S}, \alpha_0,
\beta_0, \dots, \alpha_d, \beta_d)^\intercal$.

@@

\[ \mathcal{L}(\bm\nu) = \mathbb{E}[\log\pfull] -

@@

    \mathcal{L}(\bm\nu) &= \mathbb{E}[\log p(\mathbf{X_u}) + \log p(\bm\lambda |

@@

  \mathcal{L}(\bm\nu) &= \mathbb{E}\left[-\frac{1}{2}\Tr[\Kuu^{-2}] -

@@

  \mathcal{L}(\bm\nu) &= \frac{1}{2}\log|\mathbf{S}| + \mathbb{E}\left[-\frac{1}{2}\Tr[\Kuu^{-2}] - \sum_{i=1}^d \log(\lambda_i + 1)\right] \\

@@

  \mathbb{E}[\Tr[\Kuu^{-1}\mathbf{S}] +

@@

  \Tr[\mathbb{E}[\Kuu^{-1}]\mathbf{S}] +

@@

  \mathcal{L}(\bm\nu) &= \frac{1}{2}\log|\mathbf{S}| - \frac{1}{2}\Tr[\mathbb{E}[\Kuu^{-2}]] - \sum_{i=1}^d \mathbb{E}[\log(\lambda_i + 1)] - \frac{1}{2}\Tr[\mathbb{E}[\Kuu^{-1}]\mathbf{S}] \\

@@

% TODO: finish this
Let us briefly examine the options available in case the theorem is not
applicable to some of our expected values.
\[ \frac{1}{2}\Tr[\mathbb{E}[\Kuu^{-2}]] \quad \text{and} \quad \sum_{i=1}^d
  \mathbb{E}[\log(\lambda_i + 1)] \]
come from a chosen prior distribution for $\bm\lambda | \mathbf{X_u}$.
Therefore, if needed, the distribution can be replaced with an
easier-to-deal-with alternative.

% what if the parameters of the covariance function are variational?

%mTKuum, log|Kuu|, Tr[KuuS] => p(u | lambda, X_u)
%tTKruKuum, V_r(s), T(s, a, s')V_r(s) => p(D | r) (DTC?)

We can then take the BBVI algorithm as
the basis for ours, keeping in mind the numerous improvement possibilities
provided both by the BBVI paper \cite{DBLP:conf/aistats/RanganathGB14} and many
others (e.g., the ones covered in Section \ref{literature}).

@@

The work plan can be summarised as follows:
\begin{enumerate}
\item Derive expressions for the derivatives of the ELBO w.r.t. the variational
  parameters. \emph{Deliverables}: full derivations that can be included as
  supplementary material of the final paper, including formal proofs for
  applications of the dominated convergence theorem. \emph{Date}: 15~December.
\item Implement the algorithm as a more specialised version of Algorithm~2 in
  the BBVI paper \cite{DBLP:conf/aistats/RanganathGB14}. As most of the relevant
  IRL algorithms have been implemented in
  MATLAB\footnote{\url{http://graphics.stanford.edu/projects/gpirl/irl_toolkit.zip}}\footnote{\url{https://github.com/jinming99/DGP-IRL}},
  it makes sense to do the same in order to have accurate time-sensitive
  comparisons between the algorithms. \emph{Deliverables}: a working
  implementation that can be integrated into the IRL toolkit by Levine et al.
  \cite{DBLP:conf/nips/LevinePK11}, and the same algorithm written in pseudo
  code in the paper. \emph{Date}: 1~February.
\item Evaluate the algorithm by comparing it to multiple alternatives, as
  detailed in Section \ref{sec:evaluation}. \emph{Deliverables}: multiple plots
  in an SVG or PDF format, along with code that produces them. \emph{Date}:
  1~April.
\item Write/finalise the paper. \emph{Deliverable}: a 14-page paper.
  \emph{Date}: around 23~April, the deadline has not been updated yet.
\end{enumerate}

>>>>

Given this model, data $\mathcal{D}$, and inducing feature matrix
$\mathbf{X_u}$, our goal is then to find optimal values of hyperparameters
$\bm\lambda$, inducing rewards $\mathbf{u}$, and the rewards for all relevant
states $\mathbf{r}$. While the previous paper to consider this IRL model
computed maximum likelihood estimates for $\bm\lambda$ and $\mathbf{u}$, and
made an assumption that $\mathbf{r}$ in \eqref{eq:r} has zero variance
\cite{DBLP:conf/nips/LevinePK11}, we aim to avoid this assumption and use
VI to approximate the full posterior distribution $\posterior$.
\emph{Variational inference} is an approximation technique for probability
densities \cite{blei2017variational}. Let $\approximation$ be our approximating
family of probability distributions for $\posterior$ with its own hyperparameter
vector $\bm\nu$. Then the job of VI algorithms is to optimise $\bm\nu$ in order
to minimise the \emph{Kullback-Leibler} (KL) divergence between the original
probability distribution and our approximation.  KL divergence (asymmetrically)
measures how different the two distributions are, and in this case can be
defined as \cite{blei2017variational}
\[ \begin{split}
    \DKL(\approximation || \posterior) &= \Eq [\log\approximation -
    \log\posterior ] \\
    &= \Eq[\log\approximation - \log\pfull] \\
    &+ \Eq[\log p(\mathcal{D}, \mathbf{X_u})].
  \end{split}
\]
The last term is both hard to compute and constant w.r.t. $\approximation$
\cite{blei2017variational}, so we can remove it from our optimisation objective.
The negation of what remains is often called the \emph{evidence lower bound}
(ELBO) and is defined as\footnote{Throughout the proposal, all integrals should
  be interpreted as definite integrals over the entire sample space.}
\cite{DBLP:books/lib/Bishop07,blei2017variational}
\begin{equation} \label{eq:elbo}
  \begin{split}
    \mathcal{L} &= \Eq \left[ \log \frac{\pfull}{\approximation}
    \right] \\
    &= \iiint \approximation \log
    \frac{\pfull}{\approximation}\,d\bm\lambda\,d\mathbf{u}\,d\mathbf{r}.
  \end{split}
\end{equation}

@@

Levine et al. \cite{DBLP:conf/nips/LevinePK11} are the first to suggest a way to
learn nonlinear reward functions without harsh restrictions on the problem
domain by using GPs. We base our work primarily on their paper, and the
weaknesses we hope to address have already been covered in the previous section.
A recent extension to their work by Jin et al. \cite{DBLP:conf/uai/JinDAS17}
aims to harness the power of deep learning by using several layers of GPs,
making the model less dependent on being provided good features. They also use
VI, but with a few simplifying assumptions: deterministic training conditional
for the reward vector, and fully independent training conditional for the latent
state (see Section \ref{sec:vi_literature} for details).

@@

Table \ref{table:approximations} further summarises some of the recent and/or
influential GP approximation approaches that might be relevant to our situation.
For many of the variables in our model (including the inducing points and the
hyperparameters of the covariance function), we have three possible ways of
handling them:
\begin{itemize}
\item Provide a full Bayesian treatment by defining a prior probability
  distribution. While this option can easily prevent overfitting, the choice of
  prior can be difficult to justify.
\item Fix the value of the variable. We can then optimise the value outside of
  the main algorithm, e.g., by running the algorithm with a set of possible
  values and keeping the value that produces the best performance. This option
  can be slow if we are interested in finding the optimal value, but could be
  appropriate if the variable has little impact on the algorithm's performance.
\item Treat the variable as a variational parameter. This method will
  efficiently optimise the value of the variable, but is vulnerable to
  overfitting. Furthermore, this requires us to find the derivative of the ELBO
  w.r.t. the variable, which could be difficult.
\end{itemize}

@@

In this section we show a feasible way to apply VI to the GP IRL model. Section
\ref{sec:preliminaries} defines several previously-missing parts of the model
and provides intuition for how the approximating distribution could be
structured. Section \ref{sec:structure} then defines the approximating
distribution, where some choices are justified by a body of literature, and some
are novel. Next, Section \ref{sec:elbo} derives a simplified expression for the
ELBO. Finally, in Section \ref{vi_algs} we describe our strategy for
optimising the ELBO.

@@

In order to properly investigate the difference between variational inference
and maximum likelihood estimation, we keep other parts of the model the same.
Namely, we set the covariance function to a version of the automatic relevance
detection kernel \cite{DBLP:conf/nips/LevinePK11,neal2012bayesian}
\[ k_{\bm\lambda}(\mathbf{x}_i, \mathbf{x}_j) = \lambda_0\exp\left(
    -\frac{1}{2}(\mathbf{x}_i - \mathbf{x}_j)^\intercal\bm\Lambda(\mathbf{x}_i -
    \mathbf{x}_j) - \mathbbm{1}[i \ne j]\sigma^2\Tr(\bm\Lambda) \right), \]
where $\lambda_0$ is the overall `scale' factor for how similar or distant the
states are, $\bm\Lambda = \diag(\lambda_1, \dots, \lambda_d)$ is a diagonal
matrix that determines the relevance of each feature (where $d$ denotes the
number of features), $\mathbbm{1}$ is defined as
\[ \mathbbm{1}[b] = \begin{cases}
    1 & \mbox{if $b$ is true} \\
    0 & \mbox{otherwise},
  \end{cases} \]
and $\sigma^2$ is set to $10^{-2}/2$ (as the original paper noted that
the value makes little difference to the performance of the algorithm
\cite{DBLP:conf/nips/LevinePK11}). Our vector of hyperparameters for the
covariance function is then $\bm\lambda = (\lambda_0, \dots,
\lambda_d)^\intercal$. Similarly, we keep the expression for the prior of
$\bm\lambda$:
\begin{equation} \label{eq:theta_prior}
  p(\bm\lambda | \mathbf{X_u}) = \exp\left( -\frac{1}{2}\Tr(\Kuu^{-2}) -
    \sum_{i=1}^d \log(\lambda_i + 1) \right).
\end{equation}
Defining the prior for $\bm\lambda$ means that we commit ourselves to
a Bayesian treatment of the variable as opposed to, e.g., treating it as a
variational parameter. While the remainder of the proposal will stick with this
assumption, the alternative is certainly worth exploring in its own right, and
will be investigated after the main formulation is implemented.

Next, we can rewrite the posterior by using the chain rule and Bayes' theorem in
order to get a better sense of what we are trying to approximate:
\[
  \begin{split}
    \posterior &= p(\bm\lambda | \mathbf{X_u}, \mathcal{D}) p(\mathbf{u}
    | \bm\lambda, \mathbf{X_u}, \mathcal{D}) p(\mathbf{r} | \bm\lambda,
    \mathbf{X_u}, \mathbf{u}, \mathcal{D}) \\
    &\propto p(\bm\lambda | \mathbf{X_u}, \mathcal{D}) p(\mathbf{u} |
    \bm\lambda, \mathbf{X_u}, \mathcal{D}) p(\mathcal{D} | r)
    p(\mathbf{r} | \bm\lambda, \mathbf{X_u}, \mathbf{u}) \\
    &\propto p(\bm\lambda | \mathbf{X_u}, \mathcal{D}) p(\mathcal{D} |
    \bm\lambda, \mathbf{X_u}, \mathbf{u}) p(\mathbf{u} | \bm\lambda,
    \mathbf{X_u}) p(\mathcal{D} | r) p(\mathbf{r} | \bm\lambda,
    \mathbf{X_u}, \mathbf{u}) \\
    &\propto p(\mathcal{D} | \bm\lambda, \mathbf{X_u}) p(\bm\lambda |
    \mathbf{X_u}) p(\mathcal{D} | \bm\lambda, \mathbf{X_u}, \mathbf{u})
    p(\mathbf{u} | \bm\lambda, \mathbf{X_u}) p(\mathcal{D} | r)
    p(\mathbf{r} | \bm\lambda, \mathbf{X_u}, \mathbf{u})
  \end{split}
\]
Note that now there are only two unknown probability distributions:
$p(\mathcal{D} | \bm\lambda, \mathbf{X_u})$ and $p(\mathcal{D} | \bm\lambda,
\mathbf{X_u}, \mathbf{u})$, which can be expressed as follows:
\begin{align*}
  p(\mathcal{D} | \bm\lambda, \mathbf{X_u}, \mathbf{u}) = \int &p(\mathcal{D} | r) p(\mathbf{r} | \bm\lambda, \mathbf{X_u}, \mathbf{u}) \, d\mathbf{r}, \\
  p(\mathcal{D} | \bm\lambda, \mathbf{X_u}) = \iint &p(\mathcal{D} | r) p(\mathbf{r} | \bm\lambda, \mathbf{X_u}, \mathbf{u}) p(\mathbf{u} | \bm\lambda, \mathbf{X_u}) \, d\mathbf{u} \, d\mathbf{r}.
\end{align*}
This suggests the following form for the approximation:
\begin{equation} \label{eq:approximation}
  \approximation = q(\bm\lambda) \times q(\mathbf{u} | \bm\lambda) \times q(\mathbf{r}
  | \bm\lambda, \mathbf{u}).
\end{equation}

@@

At this point, we are forced to make assumptions about the approximate posterior
in order to arrive at an implementable solution. Time permitting, ways to relax
the assumptions may be investigated towards the end of the project.

First, as is common in the literature for applying VI to GPs
\cite{DBLP:conf/nips/ChengB17,DBLP:conf/uai/HensmanFL13,DBLP:conf/icml/HoangHL15,DBLP:journals/jmlr/Titsias09},
we simply set
\begin{equation}
  q(\mathbf{r} | \bm\lambda, \mathbf{u}) = p(\mathbf{r} | \bm\lambda, \mathbf{X_u},
  \mathbf{u}).
\end{equation}
We can make a similarly justified choice for $q(\mathbf{u} | \bm\lambda)$:
\begin{equation}
  q(\mathbf{u} | \bm\lambda) = q(\mathbf{u}) = \mathcal{N}(\mathbf{u}; \mathbf{m}, \mathbf{S}),
\end{equation}
where $\mathbf{m} \in \mathbb{R}^{m}$ is the mean vector and the $m \times m$
positive semi-definite matrix $\mathbf{S}$ is the covariance matrix
\cite{DBLP:conf/nips/ChengB17,DBLP:journals/jmlr/HensmanDS17,DBLP:conf/aaai/HoangHL17}.

Next, we need to choose an approximating distribution for $\bm\lambda$, but,
unfortunately, all papers in Table \ref{table:approximations} either fix it or
treat it as a variational parameter. Hence, we make our first assumption without
justification from previous literature:
\begin{equation} \label{eq:theta}
  q(\bm\lambda) = \prod_{i=0}^dq(\lambda_i).
\end{equation}
We want to restrict $\lambda_0$ to be positive so that $k_{\bm\lambda}$ would
produce non-negative values and not become trivial. Similarly, we want that
$\lambda_i \ge 0$ for $i = 1, \dots, d$ so that $\bm\Lambda$ is a
positive-definite matrix. Considering the possible distributions for all $d + 1$
variables, we would like the mean to be flexible (i.e., not tied to zero, like
in the exponential distribution), and the tails to converge to zero as the
value of the random variable moves away from the mean. We might want to support
some right skew, but the distribution should be close to symmetric with at least
some parameter values. This limits our choice of distributions quite
significantly, and we decide to go with the gamma distribution as it is fairly
flexible and commonly used \cite{hogg2018introduction}. We then define the
probability density functions as
\begin{equation} \label{eq:zeta}
  q(\lambda_i) = \Gamma(\lambda_i; \alpha_i, \beta_i) =
  \frac{\beta_i^{\alpha_i}}{\Gamma(\alpha_i)}\lambda_i^{\alpha_i - 1}e^{-\beta_i\lambda_i}, \quad i = 0, \dots, d,
\end{equation}
where $\alpha_i > 0$ and $\beta_i > 0$ are parameters of the distribution, and
$\Gamma(\cdot)$ is the gamma function. This gives us our
vector of variational parameters $\bm\nu = (\mathbf{m}, \mathbf{S}, \alpha_0,
\beta_0, \dots, \alpha_d, \beta_d)^\intercal$.

@@

\[ \mathcal{L} = \mathbb{E}[\log\pfull] -

@@

    \mathcal{L} &= \mathbb{E}[\log p(\mathbf{X_u}) + \log p(\bm\lambda |

@@

  \mathcal{L} &= \mathbb{E}\left[-\frac{1}{2}\Tr(\Kuu^{-2}) -

@@

  \mathcal{L} &= \frac{1}{2}\log|\mathbf{S}| + \mathbb{E}\left[-\frac{1}{2}\Tr(\Kuu^{-2}) - \sum_{i=1}^d \log(\lambda_i + 1)\right] \\

@@

  \mathbb{E}[\Tr(\Kuu^{-1}\mathbf{S}) +

@@

  \Tr(\mathbb{E}[\Kuu^{-1}]\mathbf{S}) +

@@

  \mathcal{L}(\bm\nu) &= \frac{1}{2}\log|\mathbf{S}| - \frac{1}{2}\Tr(\mathbb{E}[\Kuu^{-2}]) - \sum_{i=1}^d \mathbb{E}[\log(\lambda_i + 1)] - \frac{1}{2}\Tr(\mathbb{E}[\Kuu^{-1}]\mathbf{S}) \\

@@

Let us briefly examine the options available in case the theorem is not
applicable to some of our expected values. First,
\[ \frac{1}{2}\Tr(\mathbb{E}[\Kuu^{-2}]) \quad \text{and} \quad \sum_{i=1}^d
  \mathbb{E}[\log(\lambda_i + 1)] \]
come from a chosen prior distribution for $\bm\lambda | \mathbf{X_u}$.
There are two ways to deal with them: replacing the prior with an
easier-to-handle alternative, and treating $\bm\lambda$ as a variational
parameter.

Then,
\[ \frac{1}{2}\Tr(\mathbb{E}[\Kuu^{-1}]\mathbf{S}), \quad 
  \frac{1}{2}\mathbf{m}^\intercal\mathbb{E}[\Kuu^{-1}]\mathbf{m}, \quad
  \text{and} \quad \frac{1}{2}\mathbb{E}[\log|\Kuu|] \]
all derive from $p(\mathbf{u} | \bm\lambda, \mathbf{X_u})$, which cannot be
eliminated or replaced without changing the idea of the project. However, notice
that the expressions inside $\mathbb{E}[\cdot]$ are all functions of $\Kuu$,
which depends only on $\bm\lambda$ and $\mathbf{X_u}$. Thus, if we make
$\bm\lambda$ a variational parameter, the expected values are no longer taken
w.r.t. $\bm\lambda$, and $\mathbb{E}[f(\Kuu)] = f(\Kuu)$ for any function $f$
that does not involve $\mathbf{u}$ or $\mathbf{r}$. As a result, the three
terms simplify to expressions without expected values.

Lastly,
$\mathbf{t}^\intercal\mathbb{E}[\Kru^\intercal\Kuu^{-1}]\mathbf{m}$ and both
instances of $\mathbb{E}[V_r(s)]$ come from $p(\mathcal{D} | r)$. The first
expected value can be resolved by making $\bm\lambda$ a variational
parameter. The MDP value function requires more effort. We can start by making
the deterministic training conditional assumption (as is commonly done in
previous work \cite{DBLP:conf/uai/JinDAS17,DBLP:conf/nips/LevinePK11}), which
sets the posterior covariance matrix of $\mathbf{r}$ to the zero matrix. Then
the posterior $\mathbf{r} = \Kru^\intercal\Kuu^{-1}\mathbf{u}$. Hence we still
have a (complicated) function of $\bm\lambda$ and $\mathbf{u}$ in the integral
w.r.t. both variables. The integral (or expected value) w.r.t. $\bm\lambda$ can
be eliminated the usual way, i.e., by making $\bm\lambda$ variational. As for
$\mathbf{u}$, we can borrow a trick from the deep GP IRL paper
\cite{DBLP:conf/uai/JinDAS17} and set $q(\mathbf{u}) = \delta(\mathbf{u} -
\mathbf{m})$, where $\delta$ is the Dirac delta function. The expected value
then becomes $\mathbb{E}[V_r(s)] = V_{\Kru^\intercal\Kuu^{-1}\mathbf{m}}(s)$.

We can then take the BBVI algorithm as
the basis for ours, keeping in mind the numerous improvement possibilities
provided both by the BBVI paper \cite{DBLP:conf/aistats/RanganathGB14} and many
others (see Section \ref{literature}).

@@

The work plan can be summarised as follows:
\begin{enumerate}
\item Derive expressions for the derivatives of the ELBO w.r.t. the variational
  parameters. \emph{Deliverables}: full derivations that can be included as
  supplementary material of the final paper, including formal proofs for
  applications of the dominated convergence theorem. \emph{Date}: 15~December.
\item Implement the algorithm as a more specialised version of Algorithm~2 in
  the BBVI paper \cite{DBLP:conf/aistats/RanganathGB14}. As most of the relevant
  IRL algorithms have been implemented in
  MATLAB\footnote{\url{http://graphics.stanford.edu/projects/gpirl/irl_toolkit.zip}}\footnote{\url{https://github.com/jinming99/DGP-IRL}},
  it makes sense to do the same in order to have accurate time-sensitive
  comparisons between the algorithms. \emph{Deliverables}: a working
  implementation that can be integrated into the IRL toolkit by Levine et al.
  \cite{DBLP:conf/nips/LevinePK11}, and the same algorithm written in pseudo
  code for the paper. \emph{Date}: 1~February.
\item Evaluate the algorithm by comparing it to multiple alternatives, as
  detailed in Section \ref{sec:evaluation}. \emph{Deliverables}: multiple plots
  in an SVG or PDF format, along with code that produces them. \emph{Date}:
  1~April.
\item Write/finalise the paper. \emph{Deliverable}: a 14-page paper.
  \emph{Date}: around 23~April, the deadline has not been updated yet.
\end{enumerate}
